\documentclass{article}
\usepackage{amsmath}
\usepackage{array}
\newcommand{\sign}{\operatorname{sign}}

\begin{document}

\section{B\'ezier curves with prescribed tangent directions and curvatures at the endpoints}

% <<<
Parameterization of the B\'ezier curve, for $x(t)$ and $y(t)$ equivalently,
%
\begin{align}
  x(t) &= x_0(1-t)^3 + 3x_1t(1-t)^2 + 3x_2t^2(1-t) + x_3 t^3\\
  \dot x(t) &= 3(x_1-x_0)(1-t)^2 + 3(x_3-x_2)t^2 + 6(x_2-x_1)t(1-t) \\
  \ddot x(t) &= 6(x_0 - 2x_1 + x_2)(1-t) + 6(x_1-2x_2+x_3) t \\
  \dddot x(t) &= 6(x_3-x_0) + 18(x_1-x_2)
\end{align}
%
The curvature is
\begin{equation}
  \kappa(t) = \frac{\dot x\ddot y - \ddot x\dot y}{[\dot x^2 + \dot y^2]^{3/2}}
\end{equation}
%
with the sign
%
\begin{equation}
  \begin{aligned}
    \sign\kappa &> 0 \quad\text{for left bendings}\\
    \sign\kappa &< 0 \quad\text{for right bendings.}
  \end{aligned}
\end{equation}
%
At the endpoints we can summarize:
%
\begin{align}
  \dot x(0)  &= 3(x_1 - x_0) \\
  \dot x(1)  &= 3(x_3 - x_2) \\
  \ddot x(0) &= 6(x_0 - 2x_1 + x_2) = -4\dot x(0) - 2\dot x(1) + 6(x_3 - x_0) \\
  \ddot x(1) &= 6(x_1 - 2x_2 + x_3) = +4\dot x(1) + 2\dot x(0) - 6(x_3 - x_0)
\end{align}
%
and the same for the $y$ coordinates.
The equations for the $\ddot x$ contain a convenient parameterization for this
problem, with given endpoints and tangent directions. We now introduce two
parameters for the distances between the first and the last pair of control
points:
%
\begin{align}
  \left(\begin{array}{cc}
    x_1-x_0 \\ y_1-y_0
  \end{array}\right)
  &= \frac{1}{3}
  \left(\begin{array}{cc}
    \dot x(0) \\ \dot y(0)
  \end{array}\right)
  =: \alpha
  \left(\begin{array}{cc}
    t_x(0) \\ t_y(0)
  \end{array}\right) \\
  \left(\begin{array}{cc}
    x_3-x_2 \\ y_3-y_2
  \end{array}\right)
  &= \frac{1}{3}
  \left(\begin{array}{cc}
    \dot x(1) \\ \dot y(1)
  \end{array}\right)
  =: \beta
  \left(\begin{array}{cc}
    t_x(1) \\ t_y(1)
  \end{array}\right) \qquad
\end{align}
%
Here, the externally prescribed tangent vectors are expected to be parallel to
the tangent vectors of the parameterization and normalized to 1. This implies
$\alpha>0$ and $\beta>0$ are the distances between the endpoints and their
corresponding control points.

The problem to now to find proper parameters $\alpha>0$ and $\beta>0$ for a given set of
endpoints $(x_0,y_0)$, $(x_3, y_3)$,
normalized tangent vectors $\mathbf{t}(0)$, $\mathbf{t}(1)$ and of
curvatures $\kappa(0), \kappa(1)$.
The rest of the control points is then given by
%
\begin{equation}
  x_1 = x_0 + \alpha t_x(0) \quad\text{and}\quad
  x_2 = x_3 - \beta  t_x(1).
\end{equation}
%
For the curvatures at the endpoints we get a nonlinear equation,
%
\begin{align}
  \kappa(0) (\dot x^2(0) + \dot y^2(0))^{3/2}
  &= \dot x(0) \ddot y(0) - \ddot x(0) \dot y(0) \\
  = 27 \kappa(0) |\alpha|^3
  &= \begin{aligned}[t]
      &+\dot x(0) \bigl[- 4\dot y(0) - 2\dot y(1) + 6(y_3-y_0)\bigr] \\
      &-\dot y(0) \bigl[- 4\dot x(0) - 2\dot x(1) + 6(x_3-x_0)\bigr]
     \end{aligned}\\
  &= \begin{aligned}[t]
      &-2 \bigl[\dot x(0) \dot y(1) - \dot y(0)\dot x(1)\bigr] \\
      &+6 \bigl[\dot x(0) (y_3-y_0) - \dot y(0) (x_3-x_0)\bigr]
     \end{aligned}\\
  &= \begin{aligned}[t]
      &-18 \alpha\beta \bigl[t_x(0)t_y(1) - t_y(0)t_x(1)\bigr] \\
      &+18 \alpha \bigl[t_x(0) (y_3-y_0) - t_y(0) (x_3-x_0)\bigr]
     \end{aligned}
\end{align}
%
And short,
%
\begin{equation}
  0 = \frac{3}{2}\kappa(0) \alpha^2 \sign(\alpha)
    + \beta \bigl[t_x(0)t_y(1) - t_y(0)t_x(1)\bigr]
    - \bigl[t_x(0) (y_3-y_0) - t_y(0) (x_3-x_0)\bigr]
\end{equation}
%
A similar calculation can be done for the end curvature,
%
\begin{align}
  \kappa(1) (\dot x^2(1) + \dot y^2(1))^{3/2}
  &= \dot x(1) \ddot y(1) - \ddot x(1) \dot y(1) \\
  = 27 \kappa(1) |\beta|^3
  &= \begin{aligned}[t]
       &+\dot x(1) \bigl[ 4\dot y(1) + 2\dot y(0) - 6(y_3-y_1)\bigr] \\
       &-\dot y(1) \bigl[ 4\dot x(1) + 2\dot x(0) - 6(x_3-x_1)\bigr]
     \end{aligned}\\
  &= \begin{aligned}[t]
       &-2 \bigl[\dot x(0) \dot y(1) - \dot y(0)\dot x(1)\bigr] \\
       &-6 \bigl[\dot x(1) (y_3-y_1) - \dot y(1) (x_3-x_1)\bigr]
     \end{aligned}\\
  &= \begin{aligned}[t]
       &-18 \alpha\beta \bigl[t_x(0)t_y(1) - t_y(0)t_x(1)\bigr] \\
       &-18 \beta \bigl[t_x(1) (y_3-y_1) - t_y(1) (x_3-x_1)\bigr]
     \end{aligned}
\end{align}
\begin{equation}
  0 = \frac{3}{2}\kappa(1) \beta^2 \sign(\beta)
    + \alpha \bigl[t_x(0)t_y(1) - t_y(0)t_x(1)\bigr]
    + \bigl[t_x(1) (y_3-y_1) - t_y(1) (x_3-x_1)\bigr]
\end{equation}
%
Alltogether, we find the system of equations that is to be solved.
%
\begin{gather}
  \begin{aligned}
    0 &= \frac{3}{2} \kappa(0) \alpha^2 \sign(\alpha) + \beta T - D \\
    0 &= \frac{3}{2} \kappa(1) \beta^2 \sign(\beta) + \alpha T - E
  \end{aligned}\\[\medskipamount]
  \begin{aligned}
    T &:= t_x(0)t_y(1) - t_y(0)t_x(1) \\
    D &:= \bigl[t_x(0) (y_3-y_0) - t_y(0) (x_3-x_0)\bigr]\\
    E &:= \bigl[t_x(1) (y_0-y_3) - t_y(1) (x_0-x_3)\bigr]
  \end{aligned}
\end{gather}
%
Decoupling the two equations causes problems with the absolute values,
%
\begin{align}
  \alpha = \frac{1}{T}\left(E - \frac{3}{2}\kappa(1)\beta |\beta |\right)\quad\text{and}\quad
  \beta  = \frac{1}{T}\left(D - \frac{3}{2}\kappa(0)\alpha|\alpha|\right)
\end{align}
%
Thus, for $\alpha>0$ and $\beta>0$ we need also
%
\begin{align}
  \frac{1}{T}\left(E - \frac{3}{2}\kappa(1)\beta |\beta |\right) > 0\quad\text{and}\quad
  \frac{1}{T}\left(D - \frac{3}{2}\kappa(0)\alpha|\alpha|\right) > 0
\end{align}
%
which is not always guaranteed. E.g. the combination
%
\begin{equation}
  T>0,\quad E<0\quad\text{and}\quad \kappa(1)>0
\end{equation}
%
is not compatible with a positive value of $\beta$. Under what circumstances do
we get such geometrically invalid results? Tests show that even allowing
arbitrary signs of the curvatures $\kappa(0)$~and $\kappa(1)$ does not help in
all cases.

Only when all the signs are correct, we can construct the solution by
decoupling the equations,
%
\begin{align}
  0 &= \frac{27}{8} \kappa(0)^2\kappa(1) \alpha^4
     - \frac{9}{2}D\kappa(0)\kappa(1)\alpha^2
     + T^3 \alpha
     - E T^2 + \frac{3}{2}\kappa(1) D^2 \\
   0 &= \frac{27}{8} \kappa(0)\kappa(1)^2\beta^4
     - \frac{9}{2}E\kappa(0)\kappa(1)\beta^2
     + T^3 \beta
     - D T^2 + \frac{3}{2}\kappa(0) E^2
\end{align}
%
% >>>

\section{Bounding boxes for B\'ezier curves}

% <<<
The bounding box is defined by the minimal and maximal values of a
curve. Hence the problem decouples for the two coordintes $x$ and $y$.
We can search for the minima and maxima within the valid range for the
parameter $t$ together with taking into account the values at the
boundaries at $t=0$ and $t=1$.

For the $x$ coordinate we have:
\begin{align}
  x(t) & = x_0(1-t)^3 + 3x_1t(1-t)^2 + 3x_2t^2(1-t) + x_3 t^3\\
       & = (x_3-3x_2+3x_1-x_0)t^3 + (3x_0-6x_1+3x_2)t^2 + (3x_1-3x_0)t + x_0\\
       & = a\,t^3 + \frac{3}{2}\,b\,t^2 + 3\,c\,t + x_0
\end{align}
%
The constants $a$, $b$, and $c$ are
%
\begin{align}
  a & = x_3-3x_2+3x_1-x_0 \\
  b & = 2x_0-4x_1+2x_2 \\
  c & = x_1-x_0
\end{align}
%
Now $\dot x(t)$ is
%
\begin{equation}
  \dot x(t) = 3\left[\,a\,t^2 + b\,t + c\,\right]
\end{equation}
%
For a numerically stable calculation of the roots of the function, we
first compute
%
\begin{equation}
  q = -\frac{1}{2}\left[\,b+\mathrm{sgn}(b)\sqrt{b^2-4ac}\right]\, .
\end{equation}
%
In case the square root is negative, we only need to take into account
the values at the boundaries. Otherwise we need to take into account
the two solutions
%
\begin{equation}
  t_1 = \frac{q}{a} \quad\text{and}\quad t_2 = \frac{c}{q}
\end{equation}
%
Again, for numerical failures (divisions by zero), we just need to skip
the particular solution. The minima and maxima for $x(t)$ can now
occur at $t=0$, $t=t_1$, $t=t_2$ and $t=1$.

% >>>

\end{document}

% vim:foldmethod=marker:foldmarker=<<<,>>>
